<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Spatial and spatio-temporal objects in google charts}
-->
# Spatial and spatio-temporal objects in google charts
======================================================
[Google
charts](https://developers.google.com/chart/interactive/docs/index)
are interactive graphs in web pages
that use google proprietary code. R pachage
[googleVis](http://cran.r-project.org/package=googleVis) converts
R data.frame objects into google charts. This vignette uses
[knitr](http://cran.r-project.org/package=knitr) and markdown to
create a web page from the R-markdown [source file](https://github.com/edzer/spacetime/blob/master/vignettes/stgvis.Rmd). It
was inspired by, and partly copies, the corresponding
[googleVis](http://cran.r-project.org/package=googleVis)
[vignette](http://cran.r-project.org/web/packages/googleVis/vignettes/Using_googleVis_with_knitr.html)

Set the googleVis options first to change the behaviour of `plot.gvis`,
so that only the chart component of the HTML file is written into
the output file.
```{r setOptions, message=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```
Following plot statements for `gvis` objects will automatically return
the HTML required for the 'knitted' output.

## Geo Charts
Geo charts work with country or region (state,
administrative regions, e.g. NUTS1, formally [ISO
3166-2](http://en.wikipedia.org/wiki/ISO_3166-2)) data. We will
try to make the DE_NUTS1 data in spacetime ready for this.
We read the table with names and codes, using help from
[stackoverflow](http://stackoverflow.com/questions/1395528/):
```{r results='asis'}
library(XML)
url = "http://en.wikipedia.org/wiki/ISO_3166-2:DE"
tables <- readHTMLTable(url, stringsAsFactors = FALSE)[[1]]

library(spacetime)
data(air) # loads rural and DE_NUTS1
Tbl <- gvisTable(tables, options=list(width=400))
plot(Tbl)
```
Luckily, the regions in `DE_NUTS` are in the same (alphabetical) order
as in the table downloaded from wikipedia, so we can simply copy them 
without need for matching. We will correct the two bilangual entries:
```{r}
DE_NUTS1$name = tables[,2]
DE_NUTS1$name[2] = "Bayern"        # Not: "Bayern (Bavaria)"
DE_NUTS1$name[9] = "Niedersachsen" # Not: "Niedersachsen (Lower Saxony)"
```
Plotting `Shape_Area`, a variable present for all regions in `DE_NUTS1`:
```{r GeoMapExample, results='asis', tidy=FALSE}
library(googleVis)
M = gvisGeoMap(DE_NUTS1@data, locationvar = "name", numvar = "Shape_Area",
	options=list(region="DE"))
plot(M)
```
reveals that Brandenburg is not recognized!  We will now try with
the ISO 3166-2 codes:
```{r results='asis'}
DE_NUTS1$code = tables[,1]
M = gvisGeoMap(DE_NUTS1@data, locationvar = "code", numvar = "Shape_Area",
	options=list(region="DE"))
plot(M)
```
which reveals that we now have all regions displayed. If however, we would
omit Berlin, as in
```{r results='asis'}
noBerlin = DE_NUTS1[tables[,1] != "DE-BE",]
M = gvisGeoMap(noBerlin@data, locationvar = "code", numvar = "Shape_Area",
	options=list(region="DE"))
plot(M)
```
we see that Brandenburg now includes Berlin. Google, did you miss a hole
here?
## An air quality example Geo chart
We can compute yearly average PM10 concentration over each
of the states:
```{r results='asis'}
DE_NUTS1.years = STF(DE_NUTS1, as.Date(c("2008-01-01", "2009-01-01")))
agg = aggregate(rural[,"2008::2009"], DE_NUTS1.years, mean, na.rm=TRUE)
d = agg[,1]@data # select time step one, take attr table of resulting SpatialPolygonsDataFrame object
d$code = tables[,1] # add region codes
M = gvisGeoMap(na.omit(d), locationvar = "code", numvar = "PM10",
	options=list(region="DE",height=350)) # drop NA values for Geo chart
Tbl <- gvisTable(d, options=list(height=380, width=200))
plot(gvisMerge(M, Tbl, horizontal=TRUE))
```
The white states received no value and correspond to `NA` values
in the R object `d`; they were omitted by the `na.omit` call.

## Motion Chart
The following example shows a motion chart of the `Produc` data used
in the spacetime manual, but does not convert the data. Time is simply
year (integer).
```{r MotionChartExample, results='asis', tidy=FALSE}
data("Produc", package = "plm")
M <- gvisMotionChart(Produc, "state", "year")
plot(M)
```
(Please note that the Motion Chart is only displayed when hosted on a
web server, or if placed in a directory which has been added to the 
trusted sources in the [security settings of Macromedia]
(http://www.macromedia.com/support/documentation/en/flashplayer/help/settings_manager04.html). 
See the googleVis package vignette for more details. )

# Iris Wind in timeAnnotation
```{r results='asis'}
library(sp)
library(gstat)
data("wind")
wind.loc$y = as.numeric(char2dms(as.character(wind.loc[["Latitude"]])))
wind.loc$x = as.numeric(char2dms(as.character(wind.loc[["Longitude"]])))
#coordinates(wind.loc) = ~x+y
#proj4string(wind.loc) = "+proj=longlat +datum=WGS84"
time = ISOdate(wind$year+1900, wind$month, wind$day)
wind.stack = stack(wind[,4:15])
names(wind.stack) = c("wind_speed", "station")
wind.stack$time = rep(time, 12)
M = gvisAnnotatedTimeLine(wind.stack, "time", "wind_speed", "station",
	options = list(width = "1000px", height = "500px",
		zoomStartTime=time[length(time)-365], zoomEndTime=max(time)))
plot(M)
```
# Rural PM10 data from spacetime: Annotation Chart
Annotation Charts can deal with gaps (missing values) in time series:
```{r results='asis'}
library(spacetime)
data(air)
d = as(rural[1:10,"2001"], "data.frame")
M = gvisAnnotationChart(d, "time", "PM10", "sp.ID",
	options = list(width = "1000px", height = "500px"))
		# zoomStartTime=d$time[1], zoomEndTime=d$time[1]+365))
plot(M)
```
The result is still a bit flaky; just try it for the full time series.

```{r resetOptions}
## Set options back to original options
options(op)
```
