<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Spatial and spatio-temporal objects in google charts}
-->
# Spatial and spatio-temporal objects in google charts
======================================================
[Google
charts](https://developers.google.com/chart/interactive/docs/index)
are interactive graphs in web pages
that use google proprietary code. R pachage
[googleVis](http://cran.r-project.org/package=googleVis) converts
R data.frame objects into google charts. This vignette uses
[knitr](http://cran.r-project.org/package=knitr) and markdown to
create a web page from the R-markdown [source file](...). It
was inspired by, and partly copies, the corresponding
[googleVis](http://cran.r-project.org/package=googleVis)
[vignette](http://cran.r-project.org/web/packages/googleVis/vignettes/Using_googleVis_with_knitr.html)

Set the googleVis options first to change the behaviour of `plot.gvis`,
so that only the chart component of the HTML file is written into
the output file.
```{r setOptions, message=FALSE}
library(googleVis)
op <- options(gvis.plot.tag='chart')
```
Following plot statements for `gvis` objects will automatically return
the HTML required for the 'knitted' output.

## Geo Charts
Geo charts work with country or region (state,
administrative regions, e.g. NUTS1, formally [ISO
3166-2](http://en.wikipedia.org/wiki/ISO_3166-2)) data. We will
try to make the DE_NUTS1 data in spacetime ready for this.
We read the table with names and codes, using help from
[stackoverflow](http://stackoverflow.com/questions/1395528/):
```{r results='asis'}
library(XML)
url = "http://en.wikipedia.org/wiki/ISO_3166-2:DE"
tables <- readHTMLTable(url, stringsAsFactors = FALSE)[[1]]

library(spacetime)
data(air) # loads rural and DE_NUTS1
Tbl <- gvisTable(tables, options=list(width=250))
plot(Tbl)
```
Luckily, the regions in `DE_NUTS` are in the same (alphabetical) order
as in the table downloaded from wikipedia, so we can simply copy them 
without need for matching. We will correct the two with double language:
```{r}
DE_NUTS1$name = tables[,2]
DE_NUTS1$name[2] = "Bayern"        # Not: "Bayern (Bavaria)"
DE_NUTS1$name[9] = "Niedersachsen" # Not: "Niedersachsen (Lower Saxony)"
```
Plotting `Shape_Area`, a variable present for all regions in `DE_NUTS1`:
```{r GeoMapExample, results='asis', tidy=FALSE}
library(googleVis)
M = gvisGeoMap(DE_NUTS1@data, locationvar = "name", numvar = "Shape_Area",
	options=list(region="DE"))
plot(M)
```
reveals that Brandenburg is not recognized!  We will now try with
the ISO 3166-2 codes:
```{r results='asis'}
DE_NUTS1$code = tables[,1]
M = gvisGeoMap(DE_NUTS1@data, locationvar = "code", numvar = "Shape_Area",
	options=list(region="DE"))
plot(M)
```
which reveals that we now have all regions displayed
## An air quality example Geo chart
and compute yearly average PM10 concentration over each
of the states:
```{r results='asis'}
DE_NUTS1.years = STF(DE_NUTS1, as.Date(c("2008-01-01", "2009-01-01")))
agg = aggregate(rural[,"2008::2009"], DE_NUTS1.years, mean, na.rm=TRUE)
d = agg[,1]@data
d$code = tables[,1]
M = gvisGeoMap(na.omit(d), locationvar = "code", numvar = "PM10",
	options=list(region="DE",height=350))
Tbl <- gvisTable(d, options=list(height=350, width=200))
plot(gvisMerge(M, Tbl, horizontal=TRUE))
```
reveals that the white states have no value to fill: these correspond
to `NA` values in the R object `d`.

## Motion Chart -- TODO: a sp or spacetime example
```{r MotionChartExample, results='asis', tidy=FALSE}
M <- gvisMotionChart(Fruits, 'Fruit', 'Year',
         options=list(width=400, height=350))
plot(M)
```
Please note that the Motion Chart is only displayed when hosted on a
web server, or if placed in a directory which has been added to the 
trusted sources in the [security settings of Macromedia]
(http://www.macromedia.com/support/documentation/en/flashplayer/help/settings_manager04.html). 
See the googleVis package vignette for more details. 

```{r resetOptions}
## Set options back to original options
options(op)
```
